---
title: "Moringa Module III Project"
author: "Daniel"
date: "7/3/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***Objectives***

The objective of the project is to carry out exploratory data analysis to understand what factors that can possibly make someone click on an add

***Metrics of Success***

To use various statistical techniques to find the mean, median, standard deviation, range correlation to find what factors might have led to clicking to an add.

***ADVERTISING COMPANY DATASET***

***loading the dataset***

previewing the first 6 items of our dataset

```{r, echo=TRUE}
#loading data in R environment
advertising=read.csv("http://bit.ly/IPAdvertisingData")

head(advertising)
```

```{r,echo=TRUE}
str(advertising)

```



Checking the dataset description

```{r, echo="TRUE"}
tail(advertising)
```

***checking dataset dimensions***

```{r,echo=TRUE}
dim(advertising)
```

##Data cleaning
***Checking for null values***

```{r,echo=TRUE}
colSums(is.na(advertising))

```

***checking the list of complete columns in our dataset***

```{r,echo=TRUE}
advertising[!complete.cases(advertising),]
```

thus our dataset doesn't have null values

***checking for duplicates in our dataset***

```{r,echo=TRUE}
advertising[duplicated(advertising),]

```

thus our data set did not contain duplicated data

***Checking the presence of outliers in the dataset***

```{r,echo=TRUE}
boxplot(advertising$Age ,col="red3")

```
thus no outliers on the Age column

```{r,echo=TRUE}

boxplot(advertising$Area.Income, col=" purple1")

```

```{r,echo=TRUE}
boxplot(advertising$Daily.Internet.Usage, col = "red3")
```

the column daily.internet usage didn't have any outliers

##Exploratory Data Analysis

***Measures of central tendency***

calculating the mean of our categorical columns
```{r echo=TRUE}
mean_Age=mean(advertising$Age)
mean_Age
mean_time=mean(advertising$Daily.Time.Spent.on.Site)
mean_time
mean_area_income=mean(advertising$Area.Income)
mean_area_income
mean_internet_usage=mean(advertising$Daily.Internet.Usage)
mean_internet_usage
```
calculating the median of the categorical columns

```{r,echo=TRUE}
median(advertising$Daily.Time.Spent.on.Site)
median(advertising$Age)
median(advertising$Area.Income)
median(advertising$Daily.Internet.Usage)
median(advertising$Clicked.on.Ad)

```
calculating the range
```{r,echo=TRUE}
range(advertising$Daily.Time.Spent.on.Site)
range(advertising$Age)
range(advertising$Area.Income)
range(advertising$Daily.Internet.Usage)
range(advertising$Clicked.on.Ad)
```
finding the standard devition
```{r,echo=TRUE}
sd(advertising$Daily.Time.Spent.on.Site)
sd(advertising$Age)
sd(advertising$Area.Income)
sd(advertising$Daily.Internet.Usage)
sd(advertising$Clicked.on.Ad)
```
```{r}
#checking for quantiles in each columns
quantile(advertising$Daily.Time.Spent.on.Site)
quantile(advertising$Age)
quantile(advertising$Area.Income)
quantile(advertising$Daily.Internet.Usage)
quantile(advertising$Clicked.on.Ad)

```
 

***Numerical***

```{r,echo=TRUE}
count_male<- table(advertising$Male)
barplot(count_male,main="Male frequency graph", xlab="Male",ylab= "frequency",legend=row.names(count_male),col=c("red","blue"))
```

```{r,echo=TRUE}
count_clicked<- table(advertising$Clicked.on.Ad)
barplot(count_clicked,main="Clicked.on.Ad frequency graph", xlab="Clicked.on.Ad",ylab= "frequency",legend=row.names(count_clicked),col=c("brown","red"))
```

From the figure we can see that the number of clicks on the ad were equal to the number of people who didn't click on the add



```{r,echo=TRUE}
#Frequency histogram to show daily time spend on site
hist(advertising$Daily.Time.Spent.on.Site,col=c("red"))
```

the amount of time is uniformly distributed with the least being beyond 100

```{r ,echo=TRUE}
hist(advertising$Age,col=c("blue"))
```

```{r ,echo=TRUE}
#frequency histogram to show the income distribution
hist(advertising$Area.Income,col = "brown")
```

The area.income is right skewed with most of the population ranging 50000 to 70000

```{r ,echo=TRUE}
#histogram to show the Age distribution
hist(advertising$Age,col=("maroon4"))
```

The age distribution is left skewed with most of the data around the 25 years and 40 years of age

##Bivariate analysis

```{r,echo=TRUE}
head(advertising)
```

```{r,echo=TRUE}
#scatter plot to show the relationship between age and Daily time spend on site
plot(advertising$Age, advertising$Daily.Time.Spent.on.Site )
```

the age between 25 and 45has great internet usage which reduces as it moves towards 60

```{r,echo=TRUE}
#scater plot to show the relationship between Age and area income
plot(advertising$Age, advertising$Area.Income )

```

from the plot majority of the people with the wealth lie around the age of 45 and below

```{r,echo=TRUE}
#scater plot to show the relationship between Age and Daily.iinternet usage

plot(advertising$Age, advertising$Daily.Internet.Usage, pch = 19, col = "red")

```

There is great internet usage between the age 25 to 35 where it then reduces as the age increases

```{r,echo=TRUE}

adverts <- subset (advertising, select = -c(Ad.Topic.Line,City,Country,Timestamp))

```

```{r,echo=TRUE}
#plotting a scatter plot for the adverts dataset
plot(adverts)
```

```{r,echo=TRUE}
res2 <-cor.test(advertising$Age, advertising$Daily.Internet.Usage,  method = "spearman")
res2

```

```{r,echo=TRUE}
res2 <-cor.test(advertising$Age, advertising$Clicked.on.Ad,  method = "spearman")

res2
```



```{r,echo=TRUE}
library(corrplot)

mydata.rcorr = cor(as.matrix(adverts))
mydata.rcorr
corrplot(mydata.rcorr)
```

from the analysis there is a strong positive correlation between daily.time.spent.on.net and daily internet usage. There is also a high positive correlation between Age and the number of clicks on add

```{r,echo=TRUE}

palette = colorRampPalette(c("green", "white", "red")) (20)
heatmap(x = mydata.rcorr, col = palette, symm = TRUE)

```

from the analysis there is a strong positive correlation between daily.time.spent.on.net and daily internet usage. There is also a high positive correlation between Age and the number of clicks on add

```{r,echo=TRUE}

library("PerformanceAnalytics")

chart.Correlation(mydata.rcorr, histogram=TRUE, pch=19)

```

from the analysis there is a strong positive c0rrelation between daily.time.spent.on.net and daily internet usage. There is also a high positive correlation between Age and the number of clicks on add 
##conclusion
There is a strong positive correlation between the Age of a person and clicking to an add
